### 2015-07-07 chapters/ipa_meets_unicode.tex
minor edits, additions

The variable encoding as allowed by the IPA becomes even more troublesome for
the tilde and diaeresis diacritics. In these cases, the IPA itself attaches
different semantics to the symbols above and below a base characters. The tilde
above a character (like in <ã>, using the \textsc{combining tilde} at
\uni{0303}) is used for nasalisation, while the tilde below a character (like in
<a̰>, using the \textsc{combining tilde below} at \uni{0330}) indicates creaky
voice. Likewise, the diaeresis above (like in <ä>, using the \textsc{combining
diaeresis} at \uni{0308}) is used for centralization, while the diaeresis below
a character (like in <a̤>, using the \textsc{combining diaeresis below} at
\uni{0324}) indicates breathy voice. These cases strengthen our plea to not
allow diacritics to switch position for typographic convenience.

@@

In the major revision of the IPA in 1932, affricates were represented by two consonants (e.g.\ tʃ), or when possible ligatures (ʧ), or with the tie-bar 
(t͡s). In the 1938 revision the tie-bar's semantics were broadened to indicate simultaneous articulation, e.g.\ in labial velars such as k͡p.

% Most consonants that involve two simultaneous places of articulation are written by combining two symbols with the 'tie bar' [~), for example [kp) which represents a voiceless labial-velar plosive.
% In other cases, such as [w), the sound involves two places of articulation simultaneously, which makes it inconvenient to display in the table.Ifseparate columns for all consonants with two places of articulation were provided, the size of the grid would become unmanageable.
% "Affricates can be written as digraphs, as ligatures, or with slur marks; thus ts, tS, dZ, ts, tS, dZ...

@@

The IPA includes a tie bar. The tie-bar is a convenient diacritic for visually
tokenizing input strings into chunks of phonetically salient groups, for example
to indicate affricates, doubly articulated consonants or diphthongs. This
character can be placed above or below the base characters, e.g. <\charis{t͡s}>
or <\charis{t͜s}>. IPA allows both options. The choice between the two symbols
is purely for legible rendering; there is no difference in semantics between the
two symbols. However, rendering is such a problematic issue for tie bars in
general that many linguists simply do not use them. Just looking at a few
different fonts already indicates that actually no font designer really gets the
placement right in combination with superscripts and subscripts. If really
necessary, we propose to standardize on the tie bar above the base characters,
using a \textsc{combining double inverted breve} at \uni{0361}.

@@

% the \textsc{undertie} at U+203F is used as a linking symbol to indicate the lack of a boundary, e.g.\ petit ami [pətit‿ami] `boyfriend' (French). This symbol looks like and can be confused with the tie-bar below.

@@

@@

% this section is wrong wrt IPA diacritics and needs updating @SM

>>>>

The variable encoding as allowed by the IPA becomes even more troublesome for
the tilde and diaeresis diacritics. In these cases, the IPA itself attaches
different semantics to the symbols above and below a base characters. The tilde
above a character (like in <ã>, using the \textsc{combining tilde} at
\uni{0303}) is used for nasalization, while the tilde below a character (like in
<a̰>, using the \textsc{combining tilde below} at \uni{0330}) indicates creaky
voice. Likewise, the diaeresis above (like in <ä>, using the \textsc{combining
diaeresis} at \uni{0308}) is used for centralization, while the diaeresis below
a character (like in <a̤>, using the \textsc{combining diaeresis below} at
\uni{0324}) indicates breathy voice. These cases strengthen our plea to not
allow diacritics to switch position for typographic convenience.

@@

@@

In the major revision of the IPA in 1932, affricates were represented by two consonants 
(e.g.\ <tʃ>), or ligatures <ʧ>, or with the tie-bar <t͡s>. In the 1938 revision the tie-bar's 
semantics were broadened to indicate simultaneous articulation, e.g.\ in labial velars 
such as <k͡p>. Thus the tie-bar is a convenient diacritic for visually tokenizing input 
strings into chunks of phonetically salient groups, including affricates, doubly articulated 
consonants or diphthongs.

The tie bar can be placed above or below the base characters, e.g. <\charis{t͡s}>
or <\charis{t͜s}>.\footnote{Note that the \textsc{undertie} at \uni{203F} symbol 
looks like and is easily confused with the tie-bar below. However, it is a different 
character and has a different function in IPA. The undertie is used as a linking 
symbol to indicate the lack of a boundary, e.g.\ petit ami [pətit‿ami] `boyfriend' 
($<$ French).} IPA allows both options. The choice between the two symbols
is purely for legible rendering; there is no difference in semantics between the
two symbols. However, rendering is such a problematic issue for tie bars in
general that many linguists simply do not use them. Just looking at a few
different fonts already indicates that actually no font designer really gets the
placement right in combination with superscripts and subscripts. If really
necessary, we propose to standardize on the tie bar above the base characters,
using a \textsc{combining double inverted breve} at \uni{0361}.

@@

@@

% TODO: link back to discussion on IPA principles and how ligatures got brought along historically

@@

% TODO: this section is wrong wrt IPA diacritics and needs updating @SM
